---
title: "CMS Open Payments"
author: "Benjamin Chan (chanb@ohsu.edu)"
output:
  html_document:
    keep_md: yes
    toc: yes
---

**Play around with the CMS [Open Payments](http://www.cms.gov/OpenPayments) data.**

`r Sys.Date()`

`r R.Version()$version.string`


Run the script to read the data.
Hide the code and the output since it's under version control on [GitHub](https://github.com/benjamin-chan/OpenPayments).

```{r, echo=FALSE, results="hide"}
source("getData.r")
```


Find the 10 largest payments.

```{r, results='asis'}
topPayments <- sort(D[, Total_Amount_of_Payment_USDollars], decreasing=TRUE)[1:10]
top <- D[Total_Amount_of_Payment_USDollars %in% topPayments,
         list(Total_Amount_of_Payment_USDollars,
              Date_of_Payment,
              Submitting_Applicable_Manufacturer_or_Applicable_GPO_Name,
              Physician_First_Name,
              Physician_Last_Name,
              Teaching_Hospital_Name,
              Recipient_City,
              Recipient_State,
              Recipient_Postal_Code,
              Name_of_Associated_Covered_Drug_or_Biological1,
              Form_of_Payment_or_Transfer_of_Value,
              Nature_of_Payment_or_Transfer_of_Value)][order(Total_Amount_of_Payment_USDollars, decreasing=TRUE)]
niceTable(top)
```

How many unique drugs or biologicals are represented?

```{r}
setkey(D, Name_of_Associated_Covered_Drug_or_Biological1)
length(unique(D[, Name_of_Associated_Covered_Drug_or_Biological1]))
```

List the drugs or biologicals with more than $1M in total payments.

```{r, results='asis'}
drugs <- D[,
           list(sumPayments = sum(Total_Amount_of_Payment_USDollars),
                nRecords = .N),
           Name_of_Associated_Covered_Drug_or_Biological1][order(sumPayments, decreasing=TRUE)]
niceTable(drugs[sumPayments > 1E6])
```

List the teaching hospitals with more than $1M in total payments.

```{r, results='asis'}
hosps <- D[grep("Teaching Hospital", Covered_Recipient_Type),
           list(sumPayments = sum(Total_Amount_of_Payment_USDollars),
                nRecords = .N),
           list(Teaching_Hospital_Name, Recipient_City, Recipient_State)][order(sumPayments, decreasing=TRUE)]
niceTable(hosps[sumPayments > 1E6])
```

List the physicians with more than $1M in total payments.

```{r, results='asis'}
docs <- D[grep("Physician", Covered_Recipient_Type),
          list(sumPayments = sum(Total_Amount_of_Payment_USDollars),
               nRecords = .N),
          list(Physician_Last_Name, Physician_First_Name, Recipient_City, Recipient_State)][order(sumPayments, decreasing=TRUE)]
niceTable(docs[sumPayments > 1E6])
```

List the payments to OHSU.

```{r, results='asis'}
ohsu <- D[grep("Oregon Health & Science University", Teaching_Hospital_Name),
          list(sumPayments = sum(Total_Amount_of_Payment_USDollars),
               nRecords = .N),
          list(Teaching_Hospital_Name)]
niceTable(ohsu)
```

List total payments by state.

```{r, results='asis'}
state <- D[,
           list(sumPayments = sum(Total_Amount_of_Payment_USDollars),
                nRecords = .N),
           Recipient_State]
stateNames <- data.table(Recipient_State = state.abb, state.name)
state <- merge(state, stateNames, by="Recipient_State")
niceTable(state[order(sumPayments, decreasing=TRUE)])
```

Prepare for mapping.

```{r}
require(ggplot2)
state <- state[, region := tolower(state.name)]
mapUS <- data.table(map_data("state"))
mapUS <- merge(mapUS, state, by="region")
labels <- data.table(x=state.center$x, y=state.center$y, state.abb)
labels <- labels[!(state.abb %in% c("AK", "HI"))]
setkey(mapUS, order)
```

Map.

```{r mapUSGnrl, fig.height=10, fig.width=10}
ggplot(mapUS, aes(x=long, y=lat, group=group)) +
  labs(title="General Payments") +
  annotate("text", x=-Inf, y=-Inf, label="Source: http://www.cms.gov/OpenPayments", color="black") +
  geom_polygon(aes(fill=cut_number(sumPayments / 1E6, 5))) +
  geom_path(color="gray", linestyle=2) +
  geom_text(data=labels, aes(x=x, y=y, label=state.abb, group=NULL), size=3) +
  scale_fill_brewer("$, millions", palette="Greens") +
  coord_map() +
  theme_minimal() +
  theme(axis.text=element_blank(), axis.title=element_blank(), axis.ticks=element_blank(), panel.grid.major=element_blank())
```


List the principal investigators with more than $1M in total payments.

```{r, results='asis'}
pis <- D[,
         list(sumPayments = sum(Total_Amount_of_Payment_USDollars),
              nRecords = .N),
         list(Principal_Investigator_Name1, Teaching_Hospital_Name, Recipient_City, Recipient_State)][order(sumPayments, decreasing=TRUE)]
niceTable(pis[sumPayments > 1E6])
```

List the payments to OHSU.

```{r, results='asis'}
ohsu <- D[grep("Oregon Health & Science University", Teaching_Hospital_Name),
          list(sumPayments = sum(Total_Amount_of_Payment_USDollars),
               nRecords = .N),
          list(Principal_Investigator_Name1, Teaching_Hospital_Name)][order(sumPayments, decreasing=TRUE)]
niceTable(ohsu)
```
