---
title: "CMS Open Payments"
author: "Benjamin Chan (chanb@ohsu.edu)"
output:
  html_document:
    keep_md: yes
    toc: yes
---
  
**Play around with the CMS [Open Payments](http://www.cms.gov/OpenPayments) data.**
  
`r Sys.Date()`

`r R.Version()$version.string`


Run the script to read the data.
Hide the code and the output since it's under version control on [GitHub](https://github.com/benjamin-chan/OpenPayments).
Also, source a function library.

```{r, results="hide"}
setwd(file.path("~", "GitHub Repositories", "OpenPayments"))
source("library.r")
source("getData.r")
```


# Mapping payments

**Which states received the most payments?**
Map payments at the state level.

```{r mapStates, fig.width=10}
DAgg <- D[,
          list(value = sum(Total_Amount_of_Payment_USDollars) / 1E6,
               region = Recipient_State),
          list(Recipient_State)]
require(choroplethr)
choroplethr(DAgg, 
            lod="state", 
            title="General + Research Payments")
```


# Network of payers and recipients

The network of payers and recipients is huge.
Narrow the focus to **include only the following recipients**

* OHSU
* City of Hope National Medical Center
* Dana Farber Cancer Institute
* Burkhart, Stephen

Further narrow the focus to **include payers who made more than $10,000 in payments** to those recipients.

```{r networkOHSU, fig.height=12, fig.width=12}
DAgg <- D[,
          list(sumAmt = sum(Total_Amount_of_Payment_USDollars)),
          list(payer, recipient)]
regex <- "OREGON HEALTH|CITY OF HOPE|DANA FARBER|BURKHART, STEPHEN"
DSubset <- DAgg[grepl(regex, toupper(recipient))]
DSubset <- DSubset[grep("OREGON HEALTH", toupper(recipient)), recipient := "OHSU"]
DSubset <- DSubset[grep("CITY OF HOPE", toupper(recipient)), recipient := "City of Hope"]
DSubset <- DSubset[grep("DANA FARBER", toupper(recipient)), recipient := "Dana Farber"]
DSubset <- DSubset[grep("BURKHART, STEPHEN", toupper(recipient)), recipient := "Burkhart, Stephen"]
DSubset <- DSubset[,
                   list(sumAmt = sum(sumAmt)),
                   list(payer, recipient)]
DSubset <- DSubset[sumAmt > 1E4]
DSubset <- DSubset[order(recipient, payer)]
DSubset[, list(sumAmtMillions = sum(sumAmt) / 1E6), recipient]
require(igraph)
G <- graph.data.frame(DSubset, directed=FALSE)
V(G)$size <- 0
V(G)$color <- "grey"
E(G)$width <- scale(DSubset$sumAmt, center=FALSE, scale=TRUE)
E(G)$color <- rgb(0, 128, 0, 128, maxColorValue=255)
plot(G,
     vertex.label.color="black",
     vertex.label.family="sans",
     vertex.frame.color=NA)
```