---
title: "CMS Open Payments"
author: "Benjamin Chan (chanb@ohsu.edu)"
output:
  html_document:
    keep_md: yes
    toc: yes
---
  
**Play around with the CMS [Open Payments](http://www.cms.gov/OpenPayments) data.**
  
`r Sys.time()`

`r R.Version()$version.string`


# Get the data

Run the script to read the data.
Hide the code and the output since it's under version control on [GitHub](https://github.com/benjamin-chan/OpenPayments).
Also, source a function library.

```{r, results="hide"}
setwd(file.path("~", "GitHub Repositories", "OpenPayments"))
source("library.r")
source("getData.r")
```

Use [OpenRefine](http://openrefine.org/) to clean the data.
This requires an interactive session of OpenRefine.


# Mapping payments

**Which states received the most payments?**
Map payments at the state level.

```{r mapStates, fig.width=10}
DAgg <- D[,
          list(value = sum(amount) / 1E6,
               region = Recipient_State),
          list(Recipient_State)]
require(choroplethr)
choroplethr(DAgg, 
            lod="state", 
            title="Open Payments")
```


# Network of payers and recipients


## Select recipients

The network of payers and recipients is huge.
Narrow the focus to **include only the following recipients**

* OHSU
* City of Hope National Medical Center
* Dana Farber Cancer Institute
* Burkhart, Stephen

Further narrow the focus to **include payers who made more than $10,000 in payments** to those recipients.

```{r networkOHSU, fig.height=12, fig.width=12}
DAgg <- D[,
          list(sumAmt = sum(amount)),
          list(payer, recipient)]
regex <- "OREGON HEALTH|CITY OF HOPE|DANA FARBER|BURKHART, STEPHEN"
DSubset <- DAgg[grepl(regex, toupper(recipient))]
DSubset <- DSubset[grep("OREGON HEALTH", toupper(recipient)), recipient := "OHSU"]
DSubset <- DSubset[grep("CITY OF HOPE", toupper(recipient)), recipient := "City of Hope"]
DSubset <- DSubset[grep("DANA FARBER", toupper(recipient)), recipient := "Dana Farber"]
DSubset <- DSubset[grep("BURKHART, STEPHEN", toupper(recipient)), recipient := "Burkhart, Stephen"]
DSubset <- DSubset[,
                   list(sumAmt = sum(sumAmt)),
                   list(payer, recipient)]
DSubset <- DSubset[sumAmt > 1E4]
DSubset <- DSubset[order(recipient, payer)]
DSubset[, list(sumAmtMillions = sum(sumAmt) / 1E6), recipient]
require(igraph)
G <- graph.data.frame(DSubset, directed=FALSE)
V(G)$size <- 0
V(G)$color <- "grey"
E(G)$width <- scale(DSubset$sumAmt, center=FALSE, scale=TRUE)
E(G)$color <- rgb(0, 128, 0, 128, maxColorValue=255)
plot(G,
     vertex.label.color="black",
     vertex.label.family="sans",
     vertex.frame.color=NA)
```


## Big payments

Include payment of more than $100,000.

Create a graph for the entire network.

```{r}
D[, 
  list(n = .N,
       sumAmtMillions = sum(amount) / 1E6), 
  amtCategory][, 
               list(amtCategory,
                    n,
                    pct = n / sum(n) * 100,
                    sumAmtMillions,
                    pctSumAmt = sumAmtMillions / sum(sumAmtMillions) * 100)][order(amtCategory)]
DAgg <- D[,
          list(sumAmtMillions = sum(amount) / 1E6),
          list(payer, recipient)]
DAgg <- DAgg[sumAmtMillions > 1E0,
             `:=` (sumAmtCategory = cut(sumAmtMillions, c(1E-1, 1E0, 1E1, 1E2, 1E3), include.lowest=TRUE),
                   sumAmtScaled = scale(sumAmtMillions, center=FALSE, scale=TRUE))]
DAgg[,
     list(n = .N,
          sumAmtMillions = sum(sumAmtMillions)),
     sumAmtCategory][order(sumAmtCategory)]
DAgg <- DAgg[!is.na(sumAmtCategory)]
vertices <- rbind(unique(DAgg[, list(name = payer, type = "Payer")]),
                  unique(DAgg[, list(name = recipient, type = "Recipient")]))
G <- graph.data.frame(DAgg, vertices=vertices, directed=TRUE)
V(G)$size <- 0
V(G)$color <- ifelse(V(G)$type == "Payer", "red", "black")
E(G)$color <- rgb(0, 128, 0, 128, maxColorValue=255)
E(G)$weight <- unclass(DAgg$sumAmtCategory)
E(G)$width <- unclass(DAgg$sumAmtCategory)
```

Plot the network.
Get the file at [./OpenPayments_files/figure-html/networkBigPayments.png](https://github.com/benjamin-chan/OpenPayments/blob/master/OpenPayments_files/figure-html/networkBigPayments.png) for the full effect.

```{r networkBigPayments, fig.height=30, fig.width=30}
plot(as.undirected(G),
     vertex.label.cex=0.5,
     vertex.label.color=V(G)$color,
     vertex.label.family="sans",
     vertex.frame.color=NA)
```

Write GraphML dataset for use in [Gephi](http://gephi.github.io/).

```{r}
write.graph(G, file.path(getwd(), "openpayments.graphml"), format="graphml")
```
